---
title: "Housing Market Strategy Research"
author: "Names of your Group Members"
date: "Due Date"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(tidyverse)
library(survey)
#library.packages('pander) <- uncomment me
library(pander)

## CLEAN DATA
gss <- read.csv(file = 'gss.csv')
gss_rent <- gss %>% select(total_children, own_rent, 
                           average_hours_worked, income_family, age, marital_status)
# Remove NAs
gss_rent <- gss_rent %>% 
  drop_na(total_children) %>%
  drop_na(own_rent) %>%
  drop_na(marital_status)
# Set NAs to 0 for hours worked
gss_rent <- gss_rent %>% drop_na(average_hours_worked)
# Remove those who don't know how much they work 
# and don't know if they are paying rent
gss_rent <- filter(gss_rent, average_hours_worked != "Don't know")
gss_rent <- filter(gss_rent, own_rent != "Don't know")

gss_rent <- gss_rent %>% 
  mutate(paid_cash_on_rent = 
           ifelse(own_rent=="Rented, even if no cash rent is paid", 0, 1)) 
```

*Code and data supporting this analysis is available at:* \
https://github.com/ElricL/Housing-Market-Strategy.git

## Abstract

Here is where you give a brief (one paragraph overview of your entire paper). This should include some background/introduction, some methodology, results and conclusions.

## Introduction

Here is where you should give insight into the setting and introduce the goal of the analysis. Here you can introduce ideas and basic concepts regarding the study setting and the potential model. Again, this is the introduction, so you should be explaining the importance of the work that is ahead and hopefully build some suspense for the reader. You can also highlight what will be included in the subsequent sections.

## Data

[Introduce the data, explain why it was selected. Make sure to comment on important features and highlight any potential drawbacks to the data.]

```{r, warning=FALSE, message=FALSE}
ggplot(data=gss_rent, aes(x=own_rent)) + 
      geom_bar(aes(y = (..count..)/sum(..count..))) + labs(title = "Ownership Frequency", x = "Owned or rented", y = "Frequency") +
  coord_flip() + scale_x_discrete(labels=c("Not Owned", "Owned"))
```

__Figure 1:__ We are interested in finding common characteristics from house owners and non-owners (i.e. what makes a person likely go to each bar), to build a strategy in finding new potential customers)

## Model
Model:
```{r, warning=FALSE, message=FALSE}
n=length(gss_rent$income_family)
N=36708083 - 1941873 - 2021564 - 1948681

fpc.srs = rep(N, n)

gss.design <- svydesign(id=~1, data=gss_rent, fpc=fpc.srs)

gss_logit <- svyglm(paid_cash_on_rent ~ as.factor(income_family) + total_children + age + as.factor(average_hours_worked) + as.factor(marital_status),
                    gss.design, family="binomial")



```

\pagebreak

## Results
```{r, message=FALSE, warning=FALSE, echo=FALSE}
summary(gss_logit)$coeff %>% pander
```
__Figure []:__ Key statistical summaries of the logistic regression model.

\pagebreak

## Discussion

## Weaknesses

## Next Steps

## References:
